---
title: '**Homework_Week3** <br>'
author: "Shivani"
output:
  html_document:
    toc: true
    toc_depth: 3
    toc_float: true
    number_sections: true
    date: "2024-10-07"
---

# Contents

1. Accessing and saving data
2. Tidying data
3. Visualising Data
4. More Visualisations

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

To produce a short Rmarkdown document involving the process of loading, cleaning, and visualizing the BreedingPairs data in this data set.
```{r}
#vroom-ing data to access it from git-hub
library(vroom)
Breeding_Pair <- vroom("https://raw.githubusercontent.com/ExperimentalConservation/Bioinformatics_data/master/Workshop%203/wader_data.csv")
```
## Accessing and saving data

```{r eval=FALSE, include=FALSE}
#Saving the file to my local storage as Breeding_Pair.csv
write.csv(Breeding_Pair,"C:\\Users\\fe24041\\OneDrive - University of Bristol\\Documents\\R_programming\\Week 3\\Breeding_Pair.csv", row.names=FALSE)
```

Accessing the file into R using knitr package
```{r}
#loading the knitr package
library(knitr)
Breeding_Pair<- read.csv("C:\\Users\\fe24041\\OneDrive - University of Bristol\\Documents\\R_programming\\Week 3\\Breeding_Pair.csv")

#using kable function from the knitr package to generate table
kable(Breeding_Pair)
```

## Tidying data
Checking the structure of Breeding_Pair_data and summarising the data

```{r include=FALSE}
library(tidyverse)
library(Hmisc)
```

```{r}
class(Breeding_Pair)
describe(Breeding_Pair)
```
From above analysis, looks like Year and SDMass data are intact. However some values are missing in BreedingPairs(missing 4) and AvgMass(missing 4) columns. 

Removing missing values from the data using drop_na() function
```{r}
Breeding_Pair_na <-
  #taking original data set
  Breeding_Pair %>%
  #filtering out any values equal to NA 
  drop_na(BreedingPairs, AvgMass)
kable(Breeding_Pair_na)
```

## Visualising Data

Let's visualize the trend of number of breeding pairs over the year

```{r}
#making a ggplot
ggplot(Breeding_Pair_na, aes(x=BreedingPairs, y= Year))+
  #adding a geom point
  geom_point()
```


Looking at the plot there is abnormality in the year (2998 doesn't seem right). 
Let's filter out 2998 from the year column and re-plot with new data set.

```{r}
#filtering and removing data for the year 2998
Breeding_Pair_new <- 
  Breeding_Pair_na %>%
  filter(Year != 2998) 

#plotting Breeding pair vs Year for the new filtered of data
ggplot(Breeding_Pair_new, aes(x=Year, y=BreedingPairs))+
  geom_point()
```


## More Visualizations

Breeding pair vs Year represented as a geom smooth plot.

```{r echo=TRUE}
ggplot(data = Breeding_Pair_new, aes(x = Year, y = BreedingPairs)) + 
  geom_smooth(col = "Yellow") 
```


From above plots we can see a decline in number of Breeding pairs over the years 1984 to 2016. 

Now let's consider the relationship between declining number of breedingpairs with their average mass.

```{r echo=TRUE}
P <- ggplot(data = Breeding_Pair_new, aes(x = Year, y = BreedingPairs))

#adding the graphics 
P <- P + geom_point(aes(col=AvgMass))

# view the plot
P
```


Now let's consider the trend of average mass over the years

```{r echo=TRUE}
ggplot(data = Breeding_Pair_new, aes(x = Year, y = AvgMass)) + 
  ## points
  geom_point(col = "darkgrey") +
  ## and lines! 
  geom_smooth(col = "pink")

```

From the above plot we can see there is no much variation in average mass over the period (500-550)
